#To load the train and test dataset in R, we use following command
```{r}
train <- read.csv("C:\\Users\\Sweta\\Documents\\Data\\train.csv", header=TRUE)
test <- read.csv("C:\\Users\\Sweta\\Documents\\Data\\test.csv", header=TRUE)
```

#Add new variable to the dataset: Test data has 418 observations with 10 variables while train data has 891 observations with 11 variables.
#"Survived" variable is missing in the test data. To add this, command is:
```{r}
test.Survived <- data.frame(Survived = rep("None", nrow(test)), test[,])
```

#Combine data sets: To combine test and train data set row wise, command is 
```{r}
data.combined <- rbind (train, test.Survived)
```

#to check the data type of variables in R
```{r}
str(data.combined)
```

#To change the data type for pclass from int to factor
```{r}
data.combined$Pclass <- as.factor(data.combined$Pclass)
```

#To change the data type for survived from char to factor
```{r}
data.combined$Survived <- as.factor(data.combined$Survived)
```

#To analyze the data in tabular format with respect to "Survived" variable
```{r}
table(data.combined$Survived)
```

#To see the distribution of the data with respect to "Pclass"
```{r}
table(data.combined$Pclass)
```

#For data visualization, ggplot2 has to be installed
```{r}
library(ggplot2)
```

#We make Assumption that Upper class passenger has higher chance of survival rate than others
```{r}
train$Pclass <- as.factor(train$Pclass)
ggplot(train, aes(x = Pclass, fill = factor(Survived))) + geom_bar() +  xlab("Pclass") + ylab("Total Count") +labs(fill = "Survived")
```

#Check the name variable in train data set
```{r}
head(as.character(train$Name))
```

#To check the unique name in the data.combined set
```{r}
length(unique(as.character(data.combined$Name)))
```

#Convert the data type of Name as character.
#Find the duplicate names in data.combined.
#save them in dup.names as character data type
```{r}
dup.names <- as.character(data.combined[which(duplicated(as.character(data.combined$Name))), "Name"])
```

#Extract the records from data.combined where name is in the dup.names
```{r}
data.combined[which(data.combined$Name %in% dup.names),]
```

#Find the Name which has "Miss." in the name in data.combined dataset
#store them in misses.
```{r}
library(stringr) 
misses <- data.combined[which(str_detect(data.combined$Name, "Miss.")),]
misses[1:5,]
```

#Find the Name which has "Mrs." in the name in data.combined
#store them in mrses.
```{r}
mrses <- data.combined[which(str_detect(data.combined$Name, "Mrs.")), ]
mrses[1:5,]
```
#Store all records in "males" data frame where sex is male 
```{r}
males <- data.combined[which(train$Sex == "male"), ]
males[1:5,]
```
#Add "Title" variable in the data.combined. 
#Title contains the title of passenger such as Miss., Mrs. Etc. 
#We create a function "extractTitle" which will extract the titles from all observations. 
#These values will be stored in "Title" variable.
```{r}
extractTitle <- function(Name) {
  name <- as.character(Name)  
  if (length(grep("Miss.", Name)) > 0) {
    return ("Miss.")
  } else if (length(grep("Master.", Name)) > 0) {
    return ("Master.")
  } else if (length(grep("Mrs.", Name)) > 0) {
    return ("Mrs.")
  } else if (length(grep("Mr.", Name)) > 0) {
    return ("Mr.")
  } else {
    return ("Other")
  }
}

titles <- NULL
for (i in 1:nrow(data.combined)) {
  titles <- c(titles, extractTitle(data.combined[i,"Name"]))
}
data.combined$title <- as.factor(titles)

ggplot(data.combined[1:891,], aes(x = title, fill = Survived)) +
  geom_bar() +
  facet_wrap(~Pclass) + 
  ggtitle("Pclass") +
  xlab("Title") +
  ylab("Total Count") +
  labs(fill = "Survived")
```